---
title: "UDFSN Reproduction"
output: pdf_document
date: Sys.Date()
header-includes:
      - \usepackage{booktabs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(scales)
library(fixest)
library(knitr)
library(tidycensus)
library(purrr)
library(httr)
library(lubridate) 

```

# Introduction

The following is R code used to replicate visualizations related to the manuscript entitled "Upzoning, Density, and the Financialization of Space in Nashville" by Nicholas Forster-Benson.

# Data Sourcing

NDP (2025). Annual residential report 2025. Technical report, Nashville Downtown Partnership. <https://ctycms.com/tn-nashville/docs/2025-residential-report-1.pdf>.

U.S. Census Bureau (2000a). 2000 census tract and block numbering area definitions. Census tracts 146-147 (FIPS: 47037014600, 47037014700) and adjacent tracts in Davidson County, TN.

U.S. Census Bureau (2000b). Census 2000 summary file 3 (sf3): Median gross rent and median home value. Tables H063001 (Median Gross Rent) and H085001 (Median Home Value); Table P053001 (Median Household Income in 1999).

U.S. Census Bureau (2010). 2010 census tract and block numbering area definitions. Census tract 195 (FIPS: 47037019500) and adjacent tracts in Davidson County, TN.

U.S. Census Bureau (2022). American community survey 5-year estimates (2005-2022). Table B25064 (Median Gross Rent); Table B25077 (Median Value - Owner-Occupied Housing Units); Table B19013 (Median Household Income in the Past 12 Months).

U.S. Department of Housing and Urban Development (2023). American housing survey: Rental housing trends. HUD User. National median rent trends used for comparison baseline.

# Plots and Figures

*Loading Rent Data* (see `UDFSN Reproduction.Rmd` for code)

```{r, echo=FALSE}

# ─── Import Data ──────────────────────────────────

nash_rent <- read.csv("data/nashville_cbd_tract_rent_2000_2022.csv")
us_rent <- read.csv("data/US_rent_2000_2022.csv")



# ─── AGGREGATE TO CBD vs ADJACENT AVERAGES ──────────────────────────────────

aggregated <- nash_rent %>%
  group_by(year, category) %>%
  summarize(
    n_tracts = n(),
    estimate = mean(estimate, na.rm = TRUE),
    # For Census 2000 (no MOE), use NA; for ACS, combine MOEs
    moe = if_else(
      all(is.na(moe)),
      NA_real_,
      sqrt(sum(moe^2, na.rm = TRUE)) / n_tracts
    ),
    .groups = "drop"
  ) %>%
  mutate(
    series = case_when(
      category == "CBD" ~ "Core CBD",
      category == "Adjacent" ~ "Avg. Adjacent Tracts",
      TRUE ~ category
    )
  )



# Combine with national
rent_plot_data <- bind_rows(
  aggregated %>% select(year, estimate, moe, series),
  us_rent %>% 
    mutate(series = "U.S. National Median") %>%
    select(year, estimate, moe, series)
)

# Set factor order
rent_plot_data$series <- factor(rent_plot_data$series, levels = c(
  "Core CBD",
  "Avg. Adjacent Tracts",
  "U.S. National Median"
))


```

*Plotting Rent Data*

```{r, echo=FALSE}
# ─── PLOT ────────────────────────────────────────────────────────────────────

  p <- ggplot(rent_plot_data, aes(x = year, y = estimate,
                           color = series, linetype = series)) +

  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey45", linewidth = 0.75) +
  annotate("text", x = 2010.25, y = 2150,
           label = "Downtown Upzoning",
           hjust = 0, size = 2.9, color = "grey35", fontface = "italic") +

  geom_line(linewidth = 1.15) +

  scale_color_manual(values = c(
    "Core CBD"                  = "black",
    "Avg. Adjacent Tracts"     = "grey35",
    "U.S. National Median"      = "grey"
  )) +
  scale_linetype_manual(values = c(
    "Core CBD"                  = "solid",
    "Avg. Adjacent Tracts"     = "dotted",
    "U.S. National Median"      = "dashed"
  )) +

  scale_x_continuous(breaks = seq(2000, 2022, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(labels = label_dollar(big.mark = ","),
                     breaks = seq(400, 2400, 200),
                     expand = expansion(mult = c(0.04, 0.09))) +

  labs(
    x        = NULL,
    y        = "Median Gross Rent ($/month)",
    color    = NULL,
    linetype = NULL,
    
  ) +

  theme_classic() + #  theme_minimal(base_size = 12)
  theme(
    legend.position      = "top",
    legend.justification = "left",
    legend.key.width     = unit(1.8, "cm"),
    plot.title           = element_text(face = "bold", size = 13),
    plot.subtitle        = element_text(size = 9.5, color = "grey40",
                                        margin = margin(b = 10)),
    plot.caption         = element_text(size = 7.5, color = "grey55",
                                        hjust = 0, margin = margin(t = 10)),
    plot.margin          = margin(16, 20, 12, 16),
    panel.grid.minor     = element_blank()
  )

p


# ───  Save ──────────────────────────────────

#output_path <- "figures/nashville_rent_2000_2022.png"
#ggsave(output_path, dpi = 600)

```

Adding data for 2025 sourced from NDP (2025). Annual residential report 2025. Technical report, Nashville Downtown Partnership. <https://ctycms.com/tn-nashville/docs/2025-residential-report-1.pdf>. 

```{r, echo=FALSE}
# ───  With 2025 Data  ──────────────────────────────────


# ─── Adding Row  ──────────────────────────────────
# data from https://ctycms.com/tn-nashville/docs/2025-residential-report-1.pdf
aggregated <- aggregated %>% 
  add_row(year=2025,category="CBD"  , n_tracts=1, estimate=2280, series="Core CBD") 



# Combine with national
rent_plot_data <- bind_rows(
  aggregated %>% select(year, estimate, moe, series),
  us_rent %>% 
    mutate(series = "U.S. National Median") %>%
    select(year, estimate, moe, series)
)

# Set factor order
rent_plot_data$series <- factor(rent_plot_data$series, levels = c(
  "Core CBD",
  "Avg. Adjacent Tracts",
  "U.S. National Median"
))

# ─── Plotting   ──────────────────────────────────

p <- ggplot(rent_plot_data, aes(x = year, y = estimate,
                           color = series, linetype = series)) +

  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey45", linewidth = 0.75) +
  annotate("text", x = 2010.25, y = 2150,
           label = "Downtown Upzoning",
           hjust = 0, size = 2.9, color = "grey35", fontface = "italic") +

  geom_line(linewidth = 1.15) +

  scale_color_manual(values = c(
    "Core CBD"                  = "black",
    "Avg. Adjacent Tracts"     = "grey35",
    "U.S. National Median"      = "grey"
  )) +
  scale_linetype_manual(values = c(
    "Core CBD"                  = "solid",
    "Avg. Adjacent Tracts"     = "dotted",
    "U.S. National Median"      = "dashed"
  )) +

  scale_x_continuous(breaks = seq(2000, 2025, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(labels = label_dollar(big.mark = ","),
                     breaks = seq(400, 2400, 200),
                     expand = expansion(mult = c(0.04, 0.09))) +

  labs(
    x        = NULL,
    y        = "Median Gross Rent ($/month)",
    color    = NULL,
    linetype = NULL,
    
  ) +

  theme_classic() + #  theme_minimal(base_size = 12)
  theme(
    legend.position      = "top",
    legend.justification = "left",
    legend.key.width     = unit(1.8, "cm"),
    plot.title           = element_text(face = "bold", size = 13),
    plot.subtitle        = element_text(size = 9.5, color = "grey40",
                                        margin = margin(b = 10)),
    plot.caption         = element_text(size = 7.5, color = "grey55",
                                        hjust = 0, margin = margin(t = 10)),
    plot.margin          = margin(16, 20, 12, 16),
    panel.grid.minor     = element_blank()
  )

p


# ───  Save ──────────────────────────────────

#output_path <- "figures/nashville_rent_2000_2025.png"
#ggsave(output_path, dpi = 600)



```

*Loading Home Price Data*  (see `UDFSN Reproduction.Rmd` for code)

```{r, echo=FALSE}
# ─── Import Data ──────────────────────────────────

nash_price <- read.csv("data/nashville_cbd_tract_price_2000_2022.csv")
us_price <- read.csv("data/US_price_2000_2022.csv")


# ─── AGGREGATE TO CBD vs ADJACENT AVERAGES ──────────────────────────────────

aggregated <- nash_price %>%
  group_by(year, category) %>%
  summarize(
    n_tracts = n(),
    estimate = mean(estimate, na.rm = TRUE),
    # For Census 2000 (no MOE), use NA; for ACS, combine MOEs
    moe = if_else(
      all(is.na(moe)),
      NA_real_,
      sqrt(sum(moe^2, na.rm = TRUE)) / n_tracts
    ),
    .groups = "drop"
  ) %>%
  mutate(
    series = case_when(
      category == "CBD" ~ "Core CBD",
      category == "Adjacent" ~ "Avg. Adjacent Tracts",
      TRUE ~ category
    )
  )

# Combine with national
price_plot_data <- bind_rows(
  aggregated %>% select(year, estimate, moe, series),
  us_price %>% 
    mutate(series = "U.S. National Median") %>%
    select(year, estimate, moe, series)
)

# Set factor order
price_plot_data$series <- factor(price_plot_data$series, levels = c(
  "Core CBD",
  "Avg. Adjacent Tracts",
  "U.S. National Median"
))


```

*Ploting Home Price Data*

```{r, echo=FALSE}

# ─── PLOT ────────────────────────────────────────────────────────────────────

 p <- ggplot(price_plot_data, aes(x = year, y = estimate,
                           color = series, linetype = series)) +

  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey45", linewidth = 0.75) +
  annotate("text", x = 2010.25, y = 450000,
           label = "Downtown Upzoning",
           hjust = 0, size = 2.9, color = "grey35", fontface = "italic") +

  geom_line(linewidth = 1.15) +

  scale_color_manual(values = c(
    "Core CBD"                  = "black",
    "Avg. Adjacent Tracts"     = "grey35",
    "U.S. National Median"      = "grey"
  )) +
  scale_linetype_manual(values = c(
    "Core CBD"                  = "solid",
    "Avg. Adjacent Tracts"     = "dotted",
    "U.S. National Median"      = "dashed"
  )) +

  scale_x_continuous(breaks = seq(2000, 2025, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(labels = label_dollar(big.mark = ","),
                     breaks = seq(100000, 500000, 100000),
                     expand = expansion(mult = c(0.04, 0.09))) +

  labs(
    x        = NULL,
    y        = "Median Home Price",
    color    = NULL,
    linetype = NULL,
    
  ) +

  theme_classic() + #  theme_minimal(base_size = 12)
  theme(
    legend.position      = "top",
    legend.justification = "left",
    legend.key.width     = unit(1.8, "cm"),
    plot.title           = element_text(face = "bold", size = 13),
    plot.subtitle        = element_text(size = 9.5, color = "grey40",
                                        margin = margin(b = 10)),
    plot.caption         = element_text(size = 7.5, color = "grey55",
                                        hjust = 0, margin = margin(t = 10)),
    plot.margin          = margin(16, 20, 12, 16),
    panel.grid.minor     = element_blank()
  )
p

# ───  Save ──────────────────────────────────

#output_path <- "figures/nashville_home_price_2000_2022.png"
#ggsave(output_path, dpi = 600)

```

*Loading Income Data*  (see `UDFSN Reproduction.Rmd` for code)

```{r, echo=FALSE}
# ─── Load Data ──────────────────────────────────

nash_income <- read.csv("data/nashville_cbd_tract_income_2000_2022.csv")
us_income <- read.csv("data/US_income_2000_2022.csv")


# ─── AGGREGATE TO CBD vs ADJACENT AVERAGES ──────────────────────────────────

aggregated <- nash_income %>%
  group_by(year, category) %>%
  summarize(
    n_tracts = n(),
    estimate = mean(estimate, na.rm = TRUE),
    # For Census 2000 (no MOE), use NA; for ACS, combine MOEs
    moe = if_else(
      all(is.na(moe)),
      NA_real_,
      sqrt(sum(moe^2, na.rm = TRUE)) / n_tracts
    ),
    .groups = "drop"
  ) %>%
  mutate(
    series = case_when(
      category == "CBD" ~ "Core CBD",
      category == "Adjacent" ~ "Avg. Adjacent Tracts",
      TRUE ~ category
    )
  )

# Combine with national
income_plot_data <- bind_rows(
  aggregated %>% select(year, estimate, moe, series),
  us_income %>% 
    mutate(series = "U.S. National Median") %>%
    select(year, estimate, moe, series)
)

# Set factor order
income_plot_data$series <- factor(income_plot_data$series, levels = c(
  "Core CBD",
  "Avg. Adjacent Tracts",
  "U.S. National Median"
))
```

*Plotting Income Data* 

```{r, echo=FALSE}

# ─── PLOT ────────────────────────────────────────────────────────────────────

 p <- ggplot(income_plot_data, aes(x = year, y = estimate,
                           color = series, linetype = series)) +

  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey45", linewidth = 0.75) +
  annotate("text", x = 2010.25, y = 80000,
           label = "Downtown Upzoning",
           hjust = 0, size = 2.9, color = "grey35", fontface = "italic") +

  geom_line(linewidth = 1.15) +

  scale_color_manual(values = c(
    "Core CBD"                  = "black",
    "Avg. Adjacent Tracts"     = "grey35",
    "U.S. National Median"      = "grey"
  )) +
  scale_linetype_manual(values = c(
    "Core CBD"                  = "solid",
    "Avg. Adjacent Tracts"     = "dotted",
    "U.S. National Median"      = "dashed"
  )) +

  scale_x_continuous(breaks = seq(2000, 2025, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(labels = label_dollar(big.mark = ","),
                     breaks = seq(10000, 100000, 10000),
                     expand = expansion(mult = c(0.04, 0.09))) +

  labs(
    x        = NULL,
    y        = "Median Income",
    color    = NULL,
    linetype = NULL,
    
  ) +

  theme_classic() + #  theme_minimal(base_size = 12)
  theme(
    legend.position      = "top",
    legend.justification = "left",
    legend.key.width     = unit(1.8, "cm"),
    plot.title           = element_text(face = "bold", size = 13),
    plot.subtitle        = element_text(size = 9.5, color = "grey40",
                                        margin = margin(b = 10)),
    plot.caption         = element_text(size = 7.5, color = "grey55",
                                        hjust = 0, margin = margin(t = 10)),
    plot.margin          = margin(16, 20, 12, 16),
    panel.grid.minor     = element_blank()
  )
p


# ───  Save ──────────────────────────────────

#output_path <- "figures/nashville_income_price_2000_2022.png"
#ggsave(output_path, dpi = 600)

```

*Plotting DTC Population 2001-2025*

```{r, echo=FALSE}

downtown_pop <- tibble(
  Year = c(2001, 2009, 2014, 2020, 2025),
  Population = c(1960, 4986, 7685, 14000, 22000)
) 



p <- ggplot(downtown_pop, aes(x = Year, y = Population)) +
    # Vertical dotted line: Downtown Upzoning (2010)
  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey40", linewidth = 0.8) +
  annotate("text", x = 2010.3, y = 20000,
           label = "Downtown Upzoning",
           hjust = 0, size = 3, color = "grey30", fontface = "italic") +
  geom_line() +
  geom_point() +
  theme_classic() +#  theme_classic() +  theme_minimal
  scale_y_continuous(limits = c(1000, 23000), breaks = seq(1000, 22000, 3500))+
  labs(
    x = "Year",
    y = "Downtown Population (CBD)"
  )
p


#output_path <- "figures/downtown_pop.png"
#ggsave(output_path, dpi = 600)
```
