---
title: "UDFSN Reproduction"
output: pdf_document
date: Sys.Date()
header-includes:
      - \usepackage{booktabs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(scales)
library(fixest)
library(knitr)
library(tidycensus)
library(purrr)
library(httr)
library(lubridate) 

```




# Load Rent Data

```{r}

# ─── Import Data ──────────────────────────────────

nash_rent <- read.csv("data/nashville_cbd_tract_rent_2000_2022.csv")
us_rent <- read.csv("data/US_rent_2000_2022.csv")



# ─── AGGREGATE TO CBD vs ADJACENT AVERAGES ──────────────────────────────────

aggregated <- nash_rent %>%
  group_by(year, category) %>%
  summarize(
    n_tracts = n(),
    estimate = mean(estimate, na.rm = TRUE),
    # For Census 2000 (no MOE), use NA; for ACS, combine MOEs
    moe = if_else(
      all(is.na(moe)),
      NA_real_,
      sqrt(sum(moe^2, na.rm = TRUE)) / n_tracts
    ),
    .groups = "drop"
  ) %>%
  mutate(
    series = case_when(
      category == "CBD" ~ "Core CBD",
      category == "Adjacent" ~ "Avg. Adjacent Tracts",
      TRUE ~ category
    )
  )



# Combine with national
rent_plot_data <- bind_rows(
  aggregated %>% select(year, estimate, moe, series),
  us_rent %>% 
    mutate(series = "U.S. National Median") %>%
    select(year, estimate, moe, series)
)

# Set factor order
rent_plot_data$series <- factor(rent_plot_data$series, levels = c(
  "Core CBD",
  "Avg. Adjacent Tracts",
  "U.S. National Median"
))


```





# Rent Plot
```{r}
# ─── PLOT ────────────────────────────────────────────────────────────────────

  p <- ggplot(rent_plot_data, aes(x = year, y = estimate,
                           color = series, linetype = series)) +

  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey45", linewidth = 0.75) +
  annotate("text", x = 2010.25, y = 2150,
           label = "Downtown Upzoning",
           hjust = 0, size = 2.9, color = "grey35", fontface = "italic") +

  geom_line(linewidth = 1.15) +

  scale_color_manual(values = c(
    "Core CBD"                  = "black",
    "Avg. Adjacent Tracts"     = "grey35",
    "U.S. National Median"      = "grey"
  )) +
  scale_linetype_manual(values = c(
    "Core CBD"                  = "solid",
    "Avg. Adjacent Tracts"     = "dotted",
    "U.S. National Median"      = "dashed"
  )) +

  scale_x_continuous(breaks = seq(2000, 2022, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(labels = label_dollar(big.mark = ","),
                     breaks = seq(400, 2400, 200),
                     expand = expansion(mult = c(0.04, 0.09))) +

  labs(
    x        = NULL,
    y        = "Median Gross Rent ($/month)",
    color    = NULL,
    linetype = NULL,
    
  ) +

  theme_classic() + #  theme_minimal(base_size = 12)
  theme(
    legend.position      = "top",
    legend.justification = "left",
    legend.key.width     = unit(1.8, "cm"),
    plot.title           = element_text(face = "bold", size = 13),
    plot.subtitle        = element_text(size = 9.5, color = "grey40",
                                        margin = margin(b = 10)),
    plot.caption         = element_text(size = 7.5, color = "grey55",
                                        hjust = 0, margin = margin(t = 10)),
    plot.margin          = margin(16, 20, 12, 16),
    panel.grid.minor     = element_blank()
  )

p


# ───  Save ──────────────────────────────────

output_path <- "figures/nashville_rent_2000_2022.png"
ggsave(output_path, dpi = 600)




# ───  With 2025 Data  ──────────────────────────────────


# ─── Adding Row  ──────────────────────────────────
# data from https://ctycms.com/tn-nashville/docs/2025-residential-report-1.pdf
aggregated <- aggregated %>% 
  add_row(year=2025,category="CBD"  , n_tracts=1, estimate=2280, series="Core CBD") 



# Combine with national
rent_plot_data <- bind_rows(
  aggregated %>% select(year, estimate, moe, series),
  us_rent %>% 
    mutate(series = "U.S. National Median") %>%
    select(year, estimate, moe, series)
)

# Set factor order
rent_plot_data$series <- factor(rent_plot_data$series, levels = c(
  "Core CBD",
  "Avg. Adjacent Tracts",
  "U.S. National Median"
))

# ─── Plotting   ──────────────────────────────────

p <- ggplot(rent_plot_data, aes(x = year, y = estimate,
                           color = series, linetype = series)) +

  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey45", linewidth = 0.75) +
  annotate("text", x = 2010.25, y = 2150,
           label = "Downtown Upzoning",
           hjust = 0, size = 2.9, color = "grey35", fontface = "italic") +

  geom_line(linewidth = 1.15) +

  scale_color_manual(values = c(
    "Core CBD"                  = "black",
    "Avg. Adjacent Tracts"     = "grey35",
    "U.S. National Median"      = "grey"
  )) +
  scale_linetype_manual(values = c(
    "Core CBD"                  = "solid",
    "Avg. Adjacent Tracts"     = "dotted",
    "U.S. National Median"      = "dashed"
  )) +

  scale_x_continuous(breaks = seq(2000, 2025, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(labels = label_dollar(big.mark = ","),
                     breaks = seq(400, 2400, 200),
                     expand = expansion(mult = c(0.04, 0.09))) +

  labs(
    x        = NULL,
    y        = "Median Gross Rent ($/month)",
    color    = NULL,
    linetype = NULL,
    
  ) +

  theme_classic() + #  theme_minimal(base_size = 12)
  theme(
    legend.position      = "top",
    legend.justification = "left",
    legend.key.width     = unit(1.8, "cm"),
    plot.title           = element_text(face = "bold", size = 13),
    plot.subtitle        = element_text(size = 9.5, color = "grey40",
                                        margin = margin(b = 10)),
    plot.caption         = element_text(size = 7.5, color = "grey55",
                                        hjust = 0, margin = margin(t = 10)),
    plot.margin          = margin(16, 20, 12, 16),
    panel.grid.minor     = element_blank()
  )

p


# ───  Save ──────────────────────────────────

output_path <- "figures/nashville_rent_2000_2025.png"
ggsave(output_path, dpi = 600)



```
    title = "Median Gross Rent: Nashville Core CBD vs. Adjacent Tracts (2000–2022)",
    subtitle = "Census 2000 SF3 + ACS 5-Year Estimates (Table B25064) · Davidson County, TN",

    caption = paste0(
    2025 CBD from Annual Resifential Report
      "Sources: Census 2000 SF3 (H063), ACS 5-Year estimates 2005-2009 through 2018-2022 (B25064).\n",
      "Core CBD: Census 2000 (tracts 146, 147), Census 2010 (tract 195), Census 2020 (tracts 195.02, 195.03).\n",
      "Adjacent tracts within 2km of core CBD. Tract boundaries change at vertical dotted lines.\n",
      "Shaded bands = ACS margins of error (90% CI). Hollow points = Census 2000 (no MOE). Retrieved ", Sys.Date(), "."
    
    

# Load Price Data 

```{r}
# ─── Import Data ──────────────────────────────────

nash_price <- read.csv("data/nashville_cbd_tract_price_2000_2022.csv")
us_price <- read.csv("data/US_price_2000_2022.csv")


# ─── AGGREGATE TO CBD vs ADJACENT AVERAGES ──────────────────────────────────

aggregated <- nash_price %>%
  group_by(year, category) %>%
  summarize(
    n_tracts = n(),
    estimate = mean(estimate, na.rm = TRUE),
    # For Census 2000 (no MOE), use NA; for ACS, combine MOEs
    moe = if_else(
      all(is.na(moe)),
      NA_real_,
      sqrt(sum(moe^2, na.rm = TRUE)) / n_tracts
    ),
    .groups = "drop"
  ) %>%
  mutate(
    series = case_when(
      category == "CBD" ~ "Core CBD",
      category == "Adjacent" ~ "Avg. Adjacent Tracts",
      TRUE ~ category
    )
  )

# Combine with national
price_plot_data <- bind_rows(
  aggregated %>% select(year, estimate, moe, series),
  us_price %>% 
    mutate(series = "U.S. National Median") %>%
    select(year, estimate, moe, series)
)

# Set factor order
price_plot_data$series <- factor(price_plot_data$series, levels = c(
  "Core CBD",
  "Avg. Adjacent Tracts",
  "U.S. National Median"
))


```

# Plot Price 

```{r}

# ─── PLOT ────────────────────────────────────────────────────────────────────

 p <- ggplot(price_plot_data, aes(x = year, y = estimate,
                           color = series, linetype = series)) +

  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey45", linewidth = 0.75) +
  annotate("text", x = 2010.25, y = 450000,
           label = "Downtown Upzoning",
           hjust = 0, size = 2.9, color = "grey35", fontface = "italic") +

  geom_line(linewidth = 1.15) +

  scale_color_manual(values = c(
    "Core CBD"                  = "black",
    "Avg. Adjacent Tracts"     = "grey35",
    "U.S. National Median"      = "grey"
  )) +
  scale_linetype_manual(values = c(
    "Core CBD"                  = "solid",
    "Avg. Adjacent Tracts"     = "dotted",
    "U.S. National Median"      = "dashed"
  )) +

  scale_x_continuous(breaks = seq(2000, 2025, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(labels = label_dollar(big.mark = ","),
                     breaks = seq(100000, 500000, 100000),
                     expand = expansion(mult = c(0.04, 0.09))) +

  labs(
    x        = NULL,
    y        = "Median Home Price",
    color    = NULL,
    linetype = NULL,
    
  ) +

  theme_classic() + #  theme_minimal(base_size = 12)
  theme(
    legend.position      = "top",
    legend.justification = "left",
    legend.key.width     = unit(1.8, "cm"),
    plot.title           = element_text(face = "bold", size = 13),
    plot.subtitle        = element_text(size = 9.5, color = "grey40",
                                        margin = margin(b = 10)),
    plot.caption         = element_text(size = 7.5, color = "grey55",
                                        hjust = 0, margin = margin(t = 10)),
    plot.margin          = margin(16, 20, 12, 16),
    panel.grid.minor     = element_blank()
  )
p

# ───  Save ──────────────────────────────────

output_path <- "figures/nashville_home_price_2000_2022.png"
ggsave(output_path, dpi = 600)

```

     title = "Median Home Value: Nashville Core CBD vs. Adjacent Tracts (2000–2022)",
    subtitle = "Census 2000 SF3 + ACS 5-Year Estimates (Table B25077) · Davidson County, TN",
    
    caption = paste0(
      "Sources: Census 2000 SF3 (H085), ACS 5-Year estimates 2005-2009 through 2018-2022 (B25077).\n",
      "⚠️ NOTE: This is SELF-REPORTED median value of owner-occupied units, NOT actual sale prices.\n",
      "Core CBD: Census 2000 (tracts 146, 147), Census 2010 (tract 195), Census 2020 (tracts 195.02, 195.03).\n",
      "Adjacent tracts within 2km of core CBD. Tract boundaries change at vertical dotted lines.\n",
      "Shaded bands = ACS margins of error (90% CI). Hollow points = Census 2000 (no MOE). Retrieved ", Sys.Date(), "."
    )












# Load Income Data

```{r}
# ─── Load Data ──────────────────────────────────

nash_income <- read.csv("data/nashville_cbd_tract_income_2000_2022.csv")
us_income <- read.csv("data/US_income_2000_2022.csv")


# ─── AGGREGATE TO CBD vs ADJACENT AVERAGES ──────────────────────────────────

aggregated <- nash_income %>%
  group_by(year, category) %>%
  summarize(
    n_tracts = n(),
    estimate = mean(estimate, na.rm = TRUE),
    # For Census 2000 (no MOE), use NA; for ACS, combine MOEs
    moe = if_else(
      all(is.na(moe)),
      NA_real_,
      sqrt(sum(moe^2, na.rm = TRUE)) / n_tracts
    ),
    .groups = "drop"
  ) %>%
  mutate(
    series = case_when(
      category == "CBD" ~ "Core CBD",
      category == "Adjacent" ~ "Avg. Adjacent Tracts",
      TRUE ~ category
    )
  )

# Combine with national
income_plot_data <- bind_rows(
  aggregated %>% select(year, estimate, moe, series),
  us_income %>% 
    mutate(series = "U.S. National Median") %>%
    select(year, estimate, moe, series)
)

# Set factor order
income_plot_data$series <- factor(income_plot_data$series, levels = c(
  "Core CBD",
  "Avg. Adjacent Tracts",
  "U.S. National Median"
))
```

# Income Plot

```{r}

# ─── PLOT ────────────────────────────────────────────────────────────────────

 p <- ggplot(income_plot_data, aes(x = year, y = estimate,
                           color = series, linetype = series)) +

  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey45", linewidth = 0.75) +
  annotate("text", x = 2010.25, y = 80000,
           label = "Downtown Upzoning",
           hjust = 0, size = 2.9, color = "grey35", fontface = "italic") +

  geom_line(linewidth = 1.15) +

  scale_color_manual(values = c(
    "Core CBD"                  = "black",
    "Avg. Adjacent Tracts"     = "grey35",
    "U.S. National Median"      = "grey"
  )) +
  scale_linetype_manual(values = c(
    "Core CBD"                  = "solid",
    "Avg. Adjacent Tracts"     = "dotted",
    "U.S. National Median"      = "dashed"
  )) +

  scale_x_continuous(breaks = seq(2000, 2025, 5),
                     expand = expansion(mult = c(0.02, 0.05))) +
  scale_y_continuous(labels = label_dollar(big.mark = ","),
                     breaks = seq(10000, 100000, 10000),
                     expand = expansion(mult = c(0.04, 0.09))) +

  labs(
    x        = NULL,
    y        = "Median Income",
    color    = NULL,
    linetype = NULL,
    
  ) +

  theme_classic() + #  theme_minimal(base_size = 12)
  theme(
    legend.position      = "top",
    legend.justification = "left",
    legend.key.width     = unit(1.8, "cm"),
    plot.title           = element_text(face = "bold", size = 13),
    plot.subtitle        = element_text(size = 9.5, color = "grey40",
                                        margin = margin(b = 10)),
    plot.caption         = element_text(size = 7.5, color = "grey55",
                                        hjust = 0, margin = margin(t = 10)),
    plot.margin          = margin(16, 20, 12, 16),
    panel.grid.minor     = element_blank()
  )
p


# ───  Save ──────────────────────────────────

output_path <- "figures/nashville_income_price_2000_2022.png"
ggsave(output_path, dpi = 600)

```
    title = "Median Household Income: Nashville Core CBD vs. Adjacent Tracts (2000–2022)",
    subtitle = "Census 2000 SF3 + ACS 5-Year Estimates (Table B19013) · Davidson County, TN",
    caption = paste0(
      "Sources: Census 2000 SF3 (P053 - income in 1999), ACS 5-Year estimates 2005-2009 through 2018-2022 (B19013).\n",
      "Core CBD: Census 2000 (tracts 146, 147), Census 2010 (tract 195), Census 2020 (tracts 195.02, 195.03).\n",
      "Adjacent tracts within 2km of core CBD. Tract boundaries change at vertical dotted lines.\n",
      "ACS values are inflation-adjusted to survey year dollars. Shaded bands = ACS margins of error (90% CI).\n",
      "Hollow points = Census 2000 (no MOE). Retrieved ", Sys.Date(), "."
    )


# DTC Population over time
Data Sourced from NDP (2025). Annual residential report 2025. Technical report, Nashville Downtown Partnership. [https://ctycms.com/tn-nashville/docs/2025-residential-report-1.pdf] (https://ctycms.com/tn-nashville/docs/2025-residential-report-1.pdf).
```{r}

downtown_pop <- tibble(
  Year = c(2001, 2009, 2014, 2020, 2025),
  Population = c(1960, 4986, 7685, 14000, 22000)
) 



p <- ggplot(downtown_pop, aes(x = Year, y = Population)) +
    # Vertical dotted line: Downtown Upzoning (2010)
  geom_vline(xintercept = 2010, linetype = "dotted",
             color = "grey40", linewidth = 0.8) +
  annotate("text", x = 2010.3, y = 20000,
           label = "Downtown Upzoning",
           hjust = 0, size = 3, color = "grey30", fontface = "italic") +
  geom_line() +
  geom_point() +
  theme_classic() +#  theme_classic() +  theme_minimal
  scale_y_continuous(limits = c(1000, 23000), breaks = seq(1000, 22000, 3500))+
  labs(
    x = "Year",
    y = "Downtown Population (CBD)"
  )
p


output_path <- "figures/downtown_pop.png"
ggsave(output_path, dpi = 600)
```


